Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-11-15 methods_paper/fabl-paper.tex
NLTCS case study

\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
\author{Brian Kundinger\textsuperscript{a} \and
  Jerome Reiter\textsuperscript{a} \and 
  Rebecca C.~Steorts \textsuperscript{b}}
\date{
 \textsuperscript{a} Department of Statistical Science, Duke University \\
 \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
  \today}

@@

\bigskip
\begin{abstract}
%In this paper, we propose the \emph{fast beta linkage} (\texttt{fabl}) methods, which extends a recent Bayesian Fellegi Sunter model to allow for parallel computing. In addition, we propose using hashing techniques to hasten calculations and to reduce the overall computational complexity. To achieve a scalable model, we propose independent priors in order to create computational speeds and gains, where we relax the one-to-one matching assumption. Second, we resolve the one-to-one matching issue via a post processing step. Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model. Fourth, we explore other computational speeds up based upon storage efficient indexing. Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices, where we provide comparisons to the \text{BRL} package in CRAN.

In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space. This allows us to employ hashing techniques that hasten calculations and reduce the overall computational complexity.  In addition, we are able to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called \emph{storage efficient indexing}. Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
\end{abstract}

\noindent%
{\it Keywords:} bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing

@@

	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. In this paper, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases.

	Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, Fellegi and Sunter created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}. Although the need to compute comparisons for all record pairs initially restricted this family of methods, recent work by \cite{enamorado2019using} used innovative hashing techniques to scale the approach to handle large datasets.
	% Possible citation for bipartite matching. 
	%\citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.

@@

	Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost. \cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method was so computationally intensive it could only be used to link databases of less than 100 records. Most recently, \cite{sadinle_bayesian_2017} proposed the \emph{Beta Record Linkage} model (\texttt{BRL}), using an innovative prior over the space of bipartite matchings to strictly enforce one-to-one reqiurements throughout his Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review. Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets, and unusable when the number of pairwise record comparisons becomes too large.

	In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends the the Beta Record Linkage model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage. This allows us to (1) employ hashing techniques that hasten calculations and reduce computational costs; (2) compute the pairwise record comparisons over large datasets via parallel computing; and (3) reduce memory costs through our proposed method of \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, our proposed approach (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.

	In what follows, Section~\ref{sec:review-of_prior-work} reviews the prior work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section~\ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, derives the Gibbs sampler for posterior inference, and provides the loss function used to calculate the Bayes estimate for the bipartite matching. 
	Section~\ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake. In addition, we prove the computational complexity of \texttt{fabl}. Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and a case study of homicides from the El Salvadoran Civil War. Section~\ref{discussion} concludes our paper with a discussion of directions for future work.

@@

Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with respective sizes $n_1$ and $n_2$. Without loss of generality, denote the files such that $n_1 \geq n_2$. In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across datasets can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0 \quad \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become costly for large linkage tasks or for long posterior samplers. More compactly, bipartite matching can also be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.

@@

\subsection{Comparison Data}
\label{comparison-data}

@@

The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance. We can then propose thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: 
$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$
The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.

%In what follows we use notation that reflects that these comparison vectors can be modelled in number of ways. In practice however, this binning in beneficial because it allows us to model any type of comparison through a multinomial distribution. In a Bayesian context, Dirichlet priors for

\subsection{Fellegi-Sunter Model}
\label{fellegi-sunter}
The comparison data $\bm{\gamma}_{ij}$ is not sufficient to determine whether a record is a match or nonmatch because of errors that naturally occur in the data. This motivated \cite{fellegi_theory_1969} to consider the following likelihood ratio
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
\end{align}
as a weight to estimate if record pairs are a match or nonmatch. This ratio will be large if there is strong evidence of the pair being a match, and small otherwise.

In practice $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. Under the Fellegi Sunter model, we assume both that comparison vectors are independent given their matching status, and that the match status of the record pairs are independent. This allows us to model 
the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $p$ is the proportion of records that match, and $\bm{m}$ and $\bm{u}$ are vectors of parameters for the matching and nonmatching record pairs. These unknown parameters are often estimated using the EM algorithm \citep{Winkler1988}.

After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as nonlinks (with those such that $T_u < w_{ij} < T_m$ left undetermined). This is a problem in practice as transitive closures are violated. Since each $w_{ij}$ is calculated independently, there are often situation in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.

@@

The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs. Jaro provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, equation \ref{eqn:jaro} is the maximum likelihood estimate for bipartite matching.

@@

One criticism of using mixture models is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - \bm{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate
records when sampling \(Z_j\). This creates a dependency that makes the
sampler \emph{inherently serial}, which makes the sampler slow when working on large linkage tasks.

\subsection{Blocking}

In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden. In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to many missed matches, making this form of blocking undesirable in many situations.

In this paper, we use hashing and new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without the need for blocking. This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question. In practice, the following method can be combined with blocking for even greater computational gains, but all derivations, simulations, and case studies are presented without blocking.

@@

Recall our compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink. In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we propose independent priors for each component $Z_j$. We denote the \emph{fast beta prior} as follows:

@@

Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that nonmatches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\). We encourage the reader to integrate out $\pi$ from $Z_j|\pi$ under these two settings to gain deeper intuition about this prior.

Note that linkage in this setting is conducted at the \emph{record} level, rather than at the \emph{record pair} level as in the Fellegi Sunter model. That is, $\pi$ under \texttt{fabl} estimates the proportion of \emph{records} in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimates the proportion of \emph{record pairs} that are matches. We find $\pi$ to be more interpretable parameter than $\lambda$; especcially in the bipartite case, there are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows.  necessarily tends towards 0 as the size of the linkage task grows. Additionally, we emphasize that while the Fellegi Sunter model makes $n_1 \times n_2$ many independent matching decisions and \texttt{BRL} makes $n_2$ many dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$ many independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows us to achieve many of the accuracy gains from \texttt{BRL} while maintaining the efficiency gains possible by exploiting independence.

Importantly, this independence means that we have relaxed the one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). This behavior may be desirable in its own right; \cite{Newcombe1963} was interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable. In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in  section~\ref{bayes-estimate} to resolve illegal matchings.

@@

\hypertarget{fabl-model}{%
	\subsection{Model Specification}
	\label{model-specification}}

>>>>

\title{Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage  (fabl)}
\author{Brian Kundinger\textsuperscript{a} \and
  Jerome Reiter\textsuperscript{a} \and 
  Rebecca C.~Steorts \textsuperscript{b}}
\date{
 \textsuperscript{a} Department of Statistical Science, Duke University \\
 \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
  \today}

@@

\bigskip
\begin{abstract}
Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.  We propose and investigate a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.  Specifically, in \texttt{fabl} we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.

%In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space. This allows us to employ hashing techniques that hasten calculations and reduce the overall computational complexity.  In addition, we are able to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called \emph{storage efficient indexing}. Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
\end{abstract}

\noindent%
{\it Keywords:} record linkage, fusion, hashing, distributed computing, Fellegi Sunter

@@

	In many data analysis tasks, analysts seek to identify duplicate records across two databases. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, dalzell2018regression, tang2020}. In this paper, we consider bipartite record linkage, which merges two databases together that contain duplications across but not within the respective databases.

	Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, Fellegi and Sunter created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english, enamorado2019using}. 
	% Possible citation for bipartite matching. 
	%\citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.

@@

	Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost. \cite{larsen2005} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method was so computationally intensive it could only be used to link databases of less than 100 records. \cite{sadinle_bayesian_2017} proposed the \emph{Beta Record Linkage} model (\texttt{BRL}), using a prior over the space of bipartite matchings to strictly enforce one-to-one requirements throughout his Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeller can weight the relative importance of false positives and false negatives, and identify records pairings to be be decided through clerical review. Although it was shown to work on larger tasks than previous one-to-one methods, \texttt{BRL} becomes slow when working with larger datasets.

	In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. As suggested by \cite{heck2019}, we relax the one-to-one matching requirement of \texttt{BRL} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage. This allows us to (1) employ hashing techniques that hasten calculations and reduce computational costs, (2) compute the pairwise record comparisons over large datasets via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. We argue that even in cases where a bipartite matching is desired, \texttt{fabl} (with a simple post-processing procedure) provides accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.

	In what follows, Section~\ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section~\ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, derives the Gibbs sampler for posterior inference, and provides the loss function used to calculate the Bayes estimate for the bipartite matching. Section~\ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculatoins and the scale of linkage tasks amenable to \texttt{fabl}. Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and a case study of homicides from the El Salvadoran Civil War. Section~\ref{discussion} concludes with a discussion of directions for future work.

@@

Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with respective sizes $n_1$ and $n_2$. Without loss of generality, denote the files such that $n_1 \geq n_2$. In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across databases can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0 \quad \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become costly for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

@@

%\subsection{Comparison Vectors}
%\label{comparison-vectors}

@@

The simplest way to compare two records is to check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then can propose thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: 
$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$
The choice of $I_{f\ell}$ are application specific, which we discuss in our simulation and case study.

In practice, it is not feasible to make all-to-all record comparisons as the computational complexity is of order $O(n_1 \times n_2).$ The most common solution is to utilize blocking, which places similar records into partitions, or ``blocks,'' to reduce this computational burden \citep{steorts_comparison_2014, murray2016probabilistic}. In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations.

In this paper, we use hashing and new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking. This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question. In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.

\subsection{Fellegi-Sunter Model}
\label{fellegi-sunter}
The comparison data $\bm{\gamma}_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data. This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
\end{align}
as a weight to estimate if record pairs are a match or non-match. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.

In practice $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. Under the Fellegi Sunter model, we assume both that comparison vectors are independent given their matching status, and that the match status of the record pairs are independent. This allows us to model 
the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $p$ is the proportion of records that match, and $\bm{m}$ is the set of parameters $m_{fl} = P(\gamma_{ij}^f = l|Z_j =i)$ for the matching record pairs, and $\bm{u}$ is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs. These unknown parameters are often estimated using the EM algorithm \citep{Winkler1988}.

After estimating $\bm{m}$ and $\bm{u}$, the Fellegi Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review). This is a problem in practice as transitive closures can be violated. Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.

@@

The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi Sunter weights among matched pairs. Jaro provided no theoretical justification for using this approach, however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.

@@

One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - \bm{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow the sampler when working on large linkage tasks.

@@

Recall the compact representation of the matching structure $\bm{Z} = (Z_1, Z_2, \ldots Z_{n_2})$ with $Z_j \in \{1,2, \ldots, n_1, n_1 + j\}$, where $Z_j < n_1$ indicates a link and $Z_j = n_1 + j$ indicates a nonlink. In constrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{heck2019} and use independent priors for each component $Z_j$. However, unlike Heck Wortman who proposes a flat prior for $Z_j$, we propose a proper and flexible prior. We denote the fast beta prior as follows:

@@

Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. One choice of uninformative prior might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\). We note here that prior from \cite{heck2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.

Note that linkage in this setting is conducted at the record level, rather than at the record pair level as in the Fellegi Sunter model. That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, while $\lambda$ in Fellegi Sunter estimates the proportion of record pairs that are matches. We find $\pi$ to be more interpretable parameter than $\lambda$ in the bipartite case. There are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows. Additionally, while the Fellegi Sunter model makes $n_1 \times n_2$ independent matching decisions and \texttt{BRL} makes $n_2$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$  independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to achieve the benefits from \texttt{BRL} while making efficiency gains possible by exploiting independence.

Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). This behavior may be desirable; for example, we may be interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable \citep{Newcombe1963} . In situations when a bipartite estimate is desired, we can use the post processing procedure shown in section~\ref{bayes-estimate} to resolve many-to-one matchings.

@@

%\hypertarget{fabl-model}{%
%	\subsection{Model Specification}
%	\label{model-specification}} 
%
%As commonly done in the record linkage literature, we assume that agreement levels across features are conditionally independent. Additionally, we assume that any missing data is missing at random. Let  $m_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j = i)$ denote the probability of that matching records have level $\ell$ of disagreement in feature $f$, and $u_{f\ell} = P(\Gamma_{ij}^f = \ell \mid Z_j \neq i)$ represent the same probability for non-matches. In addition, denote $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$, $\bm{u}_f = (u_{f1}, \ldots, u_{fL_f})$, $\bm{m}= (m_1, \ldots m_F)$, $\bm{u}= (u_1, \ldots u_F)$, and $\Phi = (\bm{m}, \bm{u})$.

%The model can then be specified as follows:
With the \texttt{fabl} prior defined, we present the full model as follows:
